Nail: A practical interface generator for data formats

### 2014-04-25 paper/osdi14/eval.tex
Paper extract

\subsection{Programmer effort.}
\label{s:eval-effort}
To answer the second question, we implemented small example applications based on the above
grammars and compared code size with comparable applications written without parser generators. We
also compare our DNS implementation to the example DNS application provided with the Hammer parser generator.

 a functioning toy DNS
server.  In particular, we cloned the test DNS server from the Hammer
distribution to Nail.  Hammer ships with a toy DNS server written in 683
lines of code, excluding the Hammer framework itself, that responds to
any valid DNS query with a CNAME record to the domain ``spargelze.it''.
Most of this code is taken up with custom validators, semantic actions,
and data structure definitions, with only 52 lines of code defining the
grammar with Hammer's combinators.

Our DNS server consists of 148 lines of C code and 48 lines of Nail
grammar, and supports a custom zone file format with A, NS, MX, and
CNAME records. The same grammar is used, together with 98 lines of C,
to implement a functional toy clone of the \texttt{host} command-line
tool. However, because our grammar does not yet support DNS label
compression, the latter tool will occasionally reject valid real-world
DNS responses. Both clients have functional anti-spoofing measures.

It is hard to compare the programming effort required to implement
our toy DNS server to that of a real world DNS server, since
we have less functionality, in particular for DNS compression
and additional hint records that real-world DNS servers send.
However, the closest in functionality and intent is Dan Bernstein's
djbdns,\footnote{\url{http://cr.yp.to/djbdns.html}} which aims to be
a minimalist, highly secure DNS server. The latest release of djbdns,
including various support tools, is about 10,000 lines of C as measured by
\texttt{sloccount}.  We expect that it is possible to build a feature-par
version with Nail that is an order of magnitude smaller and intend to
do so.

\paragraph{Security}

\paragraph{Other issues.}

As Nail is work in progress, many parts of the implementation, syntax and
design are not complete yet and we do not yet have meaningful performance
or security metrics.

>>>>

\subsection{Programmer effort.}

%code size unzip 6.0
%2821 lines
\label{s:eval-effort}
To answer the second question, we implemented small example applications based on the above
grammars and compared code size with comparable applications that process data manually. We will
also compare a toy DNS server implemented in Nail with a similar toy DNS server provided as an
example for the Hammer parser generator. All code size measurements for C were performed by
SlocCount\cite{sloccount}. \XXX[code size is primitive, but we show orders of magnitude in the same language].

%%DNS: 
\paragraph{DNS.}

The first application we built with Nail, is a toy DNS server inspired by the example shipped with
Hammers DNS server consists of 683 lines of C, excluding the Hammer framework itself, that responds to
any valid DNS query with a CNAME record to the domain ``spargelze.it''.
Most of this code is taken up with custom validators, semantic actions,
and data structure definitions, with only 52 lines of code defining the
grammar with Hammer's combinators.

Our DNS grammar, which is partly reproduced in Figure~\ref{fig:dns-full} consists of a 48 line  Nail
grammar and 64 lines of C implementing DNS label compression. \XXX[Make this smaller]
The grammar describes both the structure of DNS packets (36 lines) and of a simple zone-file format
(16 lines) supporting A, NS, MX and CNAME records.

Our simple DNS server, which parses a zone file, listens to incoming DNS requests, parses them and
generates appropriate responses with this Nail grammar, is implemented in 149 lines of C.

 The same grammar is used, together with 98 lines of C,
to implement a functional toy clone of the \texttt{host} command-line
tool.

It is hard to compare the programming effort required to implement
our toy DNS server to that of a real world DNS server, since
we have less , in particular the additional hint records that real-world DNS servers send and a
complete lack of configuration options.
However, the closest in functionality and intent is Dan Bernstein's
djbdns,\footnote{\url{http://cr.yp.to/djbdns.html}} which aims to be
a minimalist, highly secure DNS server. The latest release of djbdns,
including various support tools, is about 10,000 lines of C as measured by
\texttt{sloccount}. \XXX[Show that we only count the DNS server parts, not the rest. List some
features we don't have]

\paragraph{ZIP.}

Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}. The ZIP
format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
security vulnerabilities in most implementations of ZIP and other archive formats. Furthermore,
Android security has been completely by-passed multiple times through parsing differentials in
handling ZIP files\cite{saurik-masterkey}. \XXX[mention talk by Julia Wolf about zip].

ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an \emph{end-of-directory
header}. This header contains a variable length comment, so it has to be located by scanning
backwards from the end of the file until a magic number and a valid length field is found. Many ZIP
implementations disagree on how to find this header in confusing situations, such as when the
comment contains the valid magic bytes\cite{wolf-berlinsides-zip}.
This end-of-directory header contains the offset and size of the \emph{ZIP directory}, which is an
array of one \emph{directory entry header} per file in the archive.
Each entry stores file meta data, such as file name, compressed and uncompressed size and a checksum,
in addition to the offset of a \emph{local file header}. The local file header duplicates most
information from the directory entry header and the compressed file contents follow it immediately.

Duplicating information made sense when ZIP files were
stored on floppy disks with slow seek times and high fault rates and memory constraints made it
impossible to keep the ZIP directory in memory or the archive was split across multiple disks.
However, care must be taken that the meta-data is consistent. For example, vulnerabilities could occur
if the length in the central directory is used to allocate memory and the length in the local
directory is used to extract without checking that they are equal first.

\XXX[condense]

We wrote a grammar for ZIP, which consists of 92 lines of Nail and 78 lines of C implementing two
stream transforms, one for the DEFLATE compression algorithm with the help of the zlib library and
one for finding the end-of-directory header. This grammar avoids the two pitfalls mentioned above:
Nail forces the programmer to specify an explicit procedure for finding the end-of-directory record
and we use dependent fields to enforce that the duplicated information is consistent.

Our grammar is also very efficient to use. Using our grammar, we build a ZIP file extractor in 50
lines of C. Because more recent versions of ZIP have added more features, such as large file support
and encryption, the closest existing tool in functionality is the historic version 5.4 of the Info-Zip unzip
utility\cite{infozip} that is shipped with most Linux distributions. The entire unzip distribution
is about 46.000 lines of code, which is mostly optimised implementation of various compression
algorithms and other configuration and portability code.

However, unzip isolates the equivalent of our Nail tool in the file extract.c, which parses the ZIP
metadata and calls various decompression routines in other files. This file measures over 1,600
lines of C.

\paragraph{Network stack.}
Write up this stack.
\subsection{Security.} 
Fuzz DNS. Fuzz ZIP. Design argument that security is hard to show.

\subsection{Performance.}

Compare DNS requests per second of BIND, DJBDNS and NailDNS on a small zonefile.
Compare extraction performance for ZIP.

### 2014-04-25 paper/osdi14/relwk.tex
we should remember to mention the workshop paper in the osdi submission
Paper extract

@@

The second class is logic errors, where application code misinterprets
input data.  This can lead to serious security consequences when two
systems disagree on the meaning of a network packet or a signed message,
as in iOS\footnote{The XNU kernel and the user code-signing verifier
interpret executable metadata differently, so the code signature
sees different bytes at a virtual address than the executable that
runs.}~\cite{evaders6} and Android\footnote{Android applications
are distributed as .zip files. Signatures are verified with a Java
program, but the program is extracted with a C program.  The Java
program interprets all fields as signed, whereas the C program
treats them as unsigned, allowing one to replace files in a signed
archive, thereby undermining Android's security model.}~\cite{saurik-masterkey} code signing and even the X.509 protocol
underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  These mistakes are
highly application-specific, and are difficult to mitigate using existing
techniques, and these mistakes can occur even in high-level languages
that guarantee memory safety.
By allowing developers to specify their data format just once, Nail
avoids logic errors and inconsistencies in parsing and output generation.

>>>>

\XXX[previous workshop paper~\cite{bangert:nail-langsec}.]

\XXX[https://github.com/qualiancy/lotus also supports arbitrary formats,
     but requires explicit encode/decode rules.  actually kind-of like
     pack/unpack.]

@@

The second class is logic errors, where application code misinterprets
input data.  This can lead to serious security consequences when two
systems disagree on the meaning of a network packet or a signed message,
as in iOS\footnote{The XNU kernel and the user code-signing verifier
interpret executable metadata differently, so the code signature
sees different bytes at a virtual address than the executable that
runs.}~\cite{evaders6} and Android\footnote{Android applications
are distributed as .zip files. Signatures are verified with a Java
program, but the program is extracted with a C program.  The Java
program interprets all fields as signed, whereas the C program
treats them as unsigned, allowing one to replace files in a signed
archive, thereby undermining Android's security model.}~\cite{saurik-masterkey}\XXX{Duplicate, we
already mention this in the eval} code signing and even the X.509 protocol
underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  These mistakes are
highly application-specific, and are difficult to mitigate using existing
techniques, and these mistakes can occur even in high-level languages
that guarantee memory safety.
By allowing developers to specify their data format just once, Nail
avoids logic errors and inconsistencies in parsing and output generation.
